---
title: "Chapitre 3"
author: "Emulie Chhor"
date: "2023-02-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("TTR")
library("forecast")
library("ggplot2")
```

# Reading data

```{r}
births <- scan("http://robjhyndman.com/tsdldata/data/nybirths.dat")
births_ts <- ts(births, frequency = 12, start = c(1946, 1))
plot.ts(births_ts)
smooted_ts <- SMA(births_ts, n=5)

# composing time series
plot(decompose(births_ts))

# adjusted with seasonality
births_seasonal <- decompose((births_ts))$seasonal
adjusted_seasonality <- births_ts - births_seasonal
plot(adjusted_seasonality)
```


# 3.1. Simple Forecasting Methods


**Apply Simple Forecasting Methods to predict future values: average, naive, seasonal naive, drift**

```{r}
# https://otexts.com/fpp2/simple-methods.html
h=35 # forecast horizon
meanf(births_ts, h)
naive(births_ts, h)
snaive(births_ts, h)
rwf(births_ts, h, drift = TRUE)

birth_forecast <- window(births_ts, start=c(1946,1), end=c(1965,1))
autoplot(births_ts) + 
  autolayer(meanf(births_ts, h), series = "Average", PI=FALSE) + 
  autolayer(naive(births_ts, h), series = "Naive", PI=FALSE) + 
  autolayer(snaive(births_ts, h), series = "Seasonal Naive", PI=FALSE) + 
  autolayer(rwf(births_ts, h, drift = TRUE), series = "Drift", PI=FALSE)  +
  ggtitle("Forecasting births in the US (in millions)") + 
  xlab("Years") + ylab("Number of Births")

```


**Using Simple Exponential Smoothing**

```{r}

# Holt-Winters
ts_holtswinter <- HoltWinters(births_ts, beta = FALSE, gamma = FALSE)
ts_holtswinter$fitted
plot(ts_holtswinter)


```

# 3.2. Transformations and Adjustments

**Adjusting using: calendar, population, inflation**

```{r}
tmp <- cbind(Monthly=births_ts, Daily=births_ts / monthdays(births_ts))
autoplot(tmp, facets = TRUE)
autoplot(births_ts) + 
  autolayer(births_ts - decompose(births_ts)$seasonal, series = "Adjusted Seasonality") + 

```

**Adjustment using Box-Cox**

```{r}
autoplot(BoxCox(births_ts, lambda = BoxCox.lambda(births_ts)))
```


**Forecasting using Bias and Simple Back adjustments**

```{r}
autoplot(births_ts) + 
  autolayer(rwf(births_ts, drift = TRUE, lambda = 0, h=50, level = 80), series = "Simple Back Adjustment", PI=FALSE) + 
  autolayer(rwf(births_ts, drift = TRUE, lambda = 0, h=50, level = 80, biasadj = TRUE), series = "Bias Adjustment", PI=FALSE) 
```

# 3.3. Residual Diagnostics

**What do we check in residual**

We check that the residual are: 
1) Uncorrelated Residuals: if residuals are correlated, then there is information not captured by model
2) Residual mean is zero: anything except 0 means there is bias in the model
3) Constant Variance
4) Residual are normally distributed

**Example: residual Diagnostics for US births**

We want to measure the residual of a model

```{r}
# residuals for naive model
naive_model <- naive(births_ts)
autoplot(residuals(naive_model)) 
mean(na.omit(naive_model$residuals))
hist(residuals(naive_model))
ggAcf(residuals(naive_model))

# check all in one
checkresiduals(naive_model)

```




