---
title: "Chapter 2 - Time Series Graphics"
author: "Emulie Chhor"
date: "30/12/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tsibble)
library(tsibbledata)
library(ggplot2)
library(dplyr)
library(ggfortify)
library(forecast)
library(feasts)
library(fabletools)
```


## 2.1. tsibble objects

#### a) Calculate monthly percentage cost in PBS dataset using `dplyr`

```{r}
a10 <- PBS %>% 
  filter(ATC2 == 'A10') %>% 
  select(c(Month, Concession, Type, Cost)) %>% 
  summarise(TotalCost = sum(Cost)) %>% 
  mutate(Cost = TotalCost /1e6)
```

#### b) Read a csv file and convert to tsbible. Date should be in quarter.

```{r}
prison <- readr::read_csv("https://OTexts.com/fpp3/extrafiles/prison_population.csv")
prison <- prison %>% 
  mutate(Quarter = yearquarter(Date)) %>% 
  select(-c("Date")) %>% 
  as_tsibble(key = c(State, Gender, Legal, Indigenous), 
             index = Quarter)
```


## 2.2. Time Plots


#### a) Create a time plot for the number of passengers from the airport Melbourne to Syney in economy class using the 'ansett' dataset.

```{r}
mel_syd_economy <- ansett %>% 
  filter(Airports=='MEL-SYD', Class == 'Economy') %>% 
  mutate(Week = as.Date(Week)) %>% 
  as.ts()
  
ggplot2::autoplot(mel_syd_economy, col='seagreen', alpha=0.7)

```

## 2.3. Time Series Patterns

## 2.4. Seasonal Plots

#### a) Create a seasonal plot using `gg_season()`

```{r}
a10 <- PBS %>% 
  filter(ATC2 == "A10") %>%
  select(Month, Concession, Type, Cost) %>%
  summarise(TotalC = sum(Cost)) %>%
  mutate(Cost = TotalC / 1e6)


a10 %>% 
  select(c("Month", "Cost")) %>% 
  as.ts() %>% 
  ggseasonplot()
```


## 2.5. Seasonal Subseries Plots

##### a) create `subseries()` plot for a10

```{r}
a10 %>% 
  select(c("Cost", "Month")) %>% 
  as.ts() %>% 
  ggsubseriesplot()

gg_subseries(a10, Cost)
```

#### b) Extract information from `tourism` dataset using seasonal and subseries plots 

```{r}
tmp <- tourism %>% 
  group_by(State, Purpose) %>% 
  summarise(Trips=sum(Trips))

holidays <- tourism %>% 
  filter(Purpose == 'Holiday') %>% 
  group_by(State) %>% 
  summarise(Trips = sum(Trips)) %>% 
  as.ts()

autoplot(holidays)

```
We see that there is a strong seasonlity for most states, but there seems to 
be an offset is some states, namely in Northern Territory and Queensland.

```{r}
tmp <- tourism %>% 
  filter(Purpose == 'Holiday') %>% 
  group_by(State) %>% 
  summarize(Trips = sum(Trips)) 

ggseasonplot(tmp, date = "Quarter", group = 'State',  value = 'Trips')
gg_season(tmp, Trips)
```


#### c) Draw subseries for holidays

```{r}
holidays <- tourism %>% 
  filter(Purpose == 'Holiday') %>% 
  group_by(State) %>% 
  summarise(Trips = sum(Trips))

holidays %>% 
  gg_subseries(Trips)
```


## 2.6. Scatterplots

```{r}

```




