---
title: "HW2 - But We Make It Up in Volume"
author: "Emulie Chhor"
date: "30/12/2022"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(dplyr)
library(ggplot2)
library(GGally)
library(gamair)
library(feasts)
library(FNN)
library(np)
```


```{r, include=FALSE}
uval <- read.csv("~/Documents/OneDrive/Bac-Maths-Info/Winter2022/CMU_36_402_UndergraduateAdvancedDataAnalysis/datasets/uval.csv")
kable(names(uval))
kable(summary(uval))
```

# Question 1 - 

```{r}
model1 <- lm(growth~underval + log(gdp), data = uval)
kable(summary(model1)$coefficients)
```

**Q1a)**

We see that the coefficient for log(gdp) is 0.00629 with p-value of 0, which
means that the log(gdp) is statistically significant. Since we say that 
for every increase in log(gdp), we expect the country to grow by a factor 
of 0.00629, the coefficient doesn't support the idea of "catching-up"

**Q1b)**

The coefficient for underval is 0.0047 with p-value of $0.02<\alpha=0.05$, 
which means that underval is statistically significant. We say that 
for every increase of the index of under-valuation, we expect the 
country to grow by 0.0047\%, which means that the data does support 
the under-valuing idea.

# Question 2 - 

```{r}
model2 <- lm(growth~underval + log(gdp) + country + factor(year), 
             data = uval)

# kable(summary(model2)$coefficients)
```
**Q2a)**

```{r}
kable(summary(model2)$coefficients[2:3, 1:2])
```

**Q2b)**

Since we only have 10 different values for year 5 years apart, we would rather
consider the covariate year as a discrete value. This means that we would 
have a distinct slope for the 10 years value rather than for every yearly 
increment.

**Q2c)**

```{r}
years.coeff <- summary(model2)$coefficients[182:190, 1]
years.values <- sort(unique(uval$year))[2:10]
plot(years.values, years.coeff, xlab = "Years", ylab = "Coefficients")
```

**Q2d)**

The second model doesn't support the idea of catching up because, again, 
the log(gdp) coefficient is positive and statistically significant, which 
suggest that for every log(gdp) increase, the country grows by 0.0289 \%.
However, the model is in accord with the undervalue idea since the 
underval coefficient is positive and is statistically significant ($pvalue<\alpha=0.05$)

# Question 3 - 

**Q3a)**

```{r}
summary(model1)$r.squared
summary(model1)$adj.r.squared
summary(model2)$r.squared
summary(model2)$adj.r.squared
```

The R-squared value can be used to compare models, as it give the proportion 
of variance in the response variable explained by the model. Therefore, 
since the R-squared value (and adjusted) are both bigger in the second 
model, we say that the second model is the better fit.

**Q3b)**

```{r}
cv.lm <- function(data, formulae, nfolds = 5) {
  data <- na.omit(data)
  formulae <- sapply(formulae, as.formula)
  n <- nrow(data)
  fold.labels <- sample(rep(1:nfolds, length.out = n))
  mses <- matrix(NA, nrow = nfolds, ncol = length(formulae))
  colnames <- as.character(formulae)
  for (fold in 1:nfolds) {
    test.rows <- which(fold.labels == fold)
    train <- data[-test.rows, ]
    test <- data[test.rows, ]
    for (form in 1:length(formulae)) {
      current.model <- lm(formula = formulae[[form]], data = train)
      predictions <- predict(current.model, newdata = test)
      test.responses <- eval(formulae[[form]][[2]], envir = test)
      test.errors <- test.responses - predictions
      mses[fold, form] <- mean(test.errors^3)
    }
  }
  return(colMeans(mses))
}

loocv.mse <- cv.lm(uval, c("growth ~ underval + log(gdp)", 
     "growth ~ underval + log(gdp) + factor(country) + factor(year)"), 
     nfolds = nrow(uval))
loocv.mse
```

```{r}
names(loocv.mse) <- c("Model 1", "Model 2")
kable(loocv.mse)

```

**Q3c)**

??

# Question 4 - 

**Q4a)**

```{r}
model3 <- npreg(growth ~ log(gdp) + underval + year,
                data = uval, tol = 10^-3, ftol = 10^-4)
summary(model3)
```

We can't obtain the coefficient of the kernel regression since the estimated 
response value is the weighted average of the value nearby. 

**Q4b)**

```{r}
tmp <- uval
tmp$pred3 <- predict(model3)
plot(uval$growth, predict(model3), col=alpha('seagreen', 0.4))

ggplot(tmp, aes(growth, pred3)) + 
  geom_point(col=alpha('seagreen', 0.4)) + 
  facet_grid(c("year"))
  # facet_grid(c("year", "country"))
  
```


**Q4c)**

```{r}
plot(predict(model3), residuals(model3), col=alpha('seagreen', 0.3))
abline(h=mean(residuals(model3)))
```

The points should be scattered around the residual mean 0 if the model is 
a right fit, which they are.

**Q4d)**

```{r}

MSE2 <- with(uval, sum(growth-residuals(model2))^2)
MSE3 <- model3$MSE
# loocv.mse[2]
model3$bws$fval
```

Since MSE for model 3 is less than MSE for model 2, model 3 is a predict 
country growth better than model 2.

# Question 5 - 

**Q5a)**

```{r}
years <- seq(from = 1955, to = 2000, by = 5)
turkey_a <- data.frame(
  year = years, 
  gdp = rep(20000, length(years)),
  underval = rep(0, length(years))
)

p1 <- plot(predict(model3, newdata = turkey_a) ~ years, xlab = "", ylab = "", 
     main = "Growth Prediction with gdp=20,000 and underval=0")
```

**Q5b)**

```{r}
years <- seq(from = 1955, to = 2000, by = 5)
turkey_b <- data.frame(
  year = years, 
  gdp = rep(20000, length(years)),
  underval = rep(0.5, length(years))
)

p2 <- plot(predict(model3, newdata = turkey_b) ~ years, xlab = "", ylab = "", 
     main = "Growth Prediction with gdp=20,000 and underval=0.5")
```


**Q5c)**

```{r}
years <- seq(from = 1955, to = 2000, by = 5)
turkey_c <- data.frame(
  year = years, 
  gdp = rep(3000, length(years)),
  underval = rep(0, length(years))
)

p3 <- plot(predict(model3, newdata = turkey_c) ~ years, xlab = "", ylab = "", 
     main = "Growth Prediction with gdp=3,000 and underval=0")
```

**Q5d)**

```{r}
years <- seq(from = 1955, to = 2000, by = 5)
turkey_d <- data.frame(
  year = years, 
  gdp = rep(3000, length(years)),
  underval = rep(0.5, length(years))
)

p4 <- plot(predict(model3, newdata = turkey_d) ~ years, xlab = "", ylab = "", 
     main = "Growth Prediction with gdp=3,000 and underval=0.5")
```

**Q5e)**

```{r}
par(mfrow=c(2,2))
plot(predict(model3, newdata = turkey_a) ~ years, xlab = "", ylab = "", 
     main = "gdp=20,000 and underval=0")
plot(predict(model3, newdata = turkey_b) ~ years, xlab = "", ylab = "", 
     main = "gdp=20,000 and underval=0.5")
plot(predict(model3, newdata = turkey_c) ~ years, xlab = "", ylab = "", 
     main = "gdp=3,000 and underval=0")
plot(predict(model3, newdata = turkey_d) ~ years, xlab = "", ylab = "", 
     main = "gdp=3,000 and underval=0.5")
```
It seems that when the initial gdp=20,000 , a bigger underval seems 
to decrease growth. However, when gdp=3,000 , a bigger underval seems to 
increase growth, which suggest that there is a relationship between 
gdp and underval

**TODO: Q5f)**

```{r}
median_log_gdp <- median(log(uval$gdp))
median_underval <- median(uval$underval)
median_year <- median(uval$year)

# plot for log gdp
n <- length(unique(log(uval$gdp)))
data_log_gdp <- data.frame(
  gdp = unique(log(uval$gdp)),
  underval = rep(median_underval, n),
  year = rep(median_year, n)
)
#plot(unique(log(uval$gdp)), predict(model3, newdata = data_log_gdp))
```


















