---
title: "HW 6 -"
author: "Emulie Chhor"
date: "29/12/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(MASS)
library(xtable)
library(knitr)
library(dplyr)
library(GGally)
```

# Question 1 - ChickWeight Data

```{r include=FALSE}

names(ChickWeight)
```

## a) Plot the data

```{r}
coplot(weight ~ Time | Chick, data = ChickWeight, type='b', show.given = FALSE)
# plot(weight~ Time, data = ChickWeight)
```


## b) Extract the data corresponding to the sixth chick. 
i) Fit a linear model using time to predict weight. Check the residuals to check 
if the model fits well. If not, plot it using polynomials.

```{r}
chick6 <- ChickWeight[ChickWeight$Chick == 6, ]
model_chick6 <- lm(weight~Time, data = chick6)
summary(model_chick6)

# plotting the model
plot(chick6$Time, chick6$weight, col='orange')
abline(model_chick6, col='seagreen')

ggplot(chick6, aes(x=Time, y=weight)) + 
  geom_point() + 
  geom_smooth(method = 'lm')

# plotting the residuals
ggplot() +
  geom_point(aes(x=chick6$Time, y=residuals(model_chick6))) + 
  geom_smooth(aes(x=chick6$Time, y=residuals(model_chick6))) + 
  xlab("Time") + ylab("Residuals")
  
# plot(chick6$Time, residuals(model_chick6), xlab = "Time", ylab = "Residuals")
# abline(h=0, col='gray')
```

By plotting the residual plot, we see that a simple linear model doesn't 
fit well because the residuals seems correlated. We will try to fit a polynomial 

```{r}
# constructing polynomial model
model_poly <- lm(weight ~ Time + I(Time ^2) + I(Time ^3) + I(Time^4), data = chick6)
with(chick6, plot(Time, weight))
lines(chick6$Time, fitted(model_poly))
# summary(model_poly)

# plotting the residuals
ggplot() + 
  geom_point(aes(x=chick6$Time, y=residuals(model_poly))) + 
  geom_smooth(aes(x=chick6$Time, y=residuals(model_poly)))

```

The polynomial model is a better fit than the simple linear regression model because the residuals don't seem to be correlated


## c) Fit a linear model to predict weight from time (ignore chick and diet) using all the data
i) Fit a linear model using time to predict weight. Check the residuals to check 
if the model fits well. If not, plot it using polynomials.

```{r}
# plotting linear regression
model_all_chicks <- lm(weight~Time, data = ChickWeight)
ggplot(ChickWeight, aes(x=Time, y=weight)) + 
  geom_point(alpha = 0.2, col='seagreen') + 
  geom_smooth(method = 'lm')

# plotting residuals
plot(ChickWeight$Time, residuals(model_all_chicks))
```

It seems that the simple linear regression isn't a good fit since the model violate the homodescacity hypothesis: the residuals keeps increasing and are therefore not constant.
We will plot the data using a polynomial 

```{r}
# plot polynomial regression
model_all_chicks_poly <- lm(weight ~ Time + I(Time^2) + I(Time ^3), data = ChickWeight)
# model_all_chicks_poly <- lm(weight ~ Time + I(Time^2) , data = ChickWeight)
with(ChickWeight, plot(x=Time, y=weight))
lines(seq(0, max(ChickWeight$Time)), 
      predict(model_all_chicks_poly, newdata = data.frame(Time = seq(0, max(ChickWeight$Time)))))
# fitted(model_all_chicks_poly)
# predict(model_all_chicks_poly, newdata = data.frame(Time = seq(0, max(ChickWeight$Time))))
# summary(model_all_chicks_poly)

# plot residuals

```


## d) Repeat c) but include Diet variable (treat it as categorical variable) (TODO)

```{r}
model4 <- lm(weight~Time + I(Time^2) + factor(Diet), data = ChickWeight)
```


# Question 2 - Income

```{r}
x0 <- rep(1, 12)
x1 <- c(1,1,1,1,0,0,0,0,0,0,0,0)
x2 <- c(0,0,0,0,1,1,1,1,0,0,0,0)
x3 <- c(0,0,0,0,0,0,0,0,1,1,1,1)
Y <- c(33, 36, 35, 35, 31, 29, 31, 29, 37, 39, 36, 36)
X <- cbind(x0, x1, x2)

```


## a) Show that if we include all three of the covariates, the $X^TX$ is not invertible. Explain why

Since $X0=X1+X2+X3$, then the columns are linearly dependant which means that $det(X^TX) = 0$. Thus, the matrix is not invertible.


## b) Fit the model: $Y=\beta_0 + \beta_1 X_1 + \beta_2 X_2 + \epsilon$

```{r}
# model6 <- lm(Y~X-1)
model6 <- lm(Y~X)
tmp <- summary(model6)$coefficients
row.names(tmp) <- c("Intercept", "France", "Italy")
kable(tmp)
```



## c) Estimate the mean income of each of the three countries in terms of $\hat{\beta_0}, \hat{\beta_1}, \hat{\beta_2}$ (CHECK)

$E[Income | France ] = \beta_0 + \beta_1$
$E[Income | Italy ] = \beta_0 + \beta_2$
$E[Income | USA ] = \beta_0 $


## d) using the regression output, construct a 95 percent confidence interval for the mean income of France

The 95% confidence interval for France's is given by the 
following formula: 

$(\beta_0 + \beta_1) \pm t_{n-2}(0.025) \sqrt{Var(\beta_0 + \beta_1)}$

$Var(\beta_0 + \beta_1) = Var(\beta_0) + Var(\beta_1)+ 2Cov(\beta_0, \beta_1)$

$Cov(\beta_0, \beta_1)=\frac{1}{n-1} \sum{(x- \bar{x})(y-\bar{y})}$

```{r}
n <- nrow(ChickWeight)
beta0 <- summary(model6)$coefficients[1,1]
beta1 <- summary(model6)$coefficients[2,1]
std_beta0 <- summary(model6)$coefficients[1,2]
std_beta1 <- summary(model6)$coefficients[2,2]
cov_b0_b1 <- 1/(n-1) *sum((X - mean(X))*(Y- mean(Y)))
std_b0_b1 <- sqrt(std_beta0^2 + std_beta1^2 + 2* cov_b0_b1)


left <- (beta0 + beta1) - qt(0.025, df=n-2, lower.tail = FALSE) * std_b0_b1
right <- (beta0 + beta1) + qt(0.025, df=n-2, lower.tail = FALSE) * std_b0_b1

```

The 95% confidence interval for France income is (32.61, 36.89)

# Question 3 - birthwt

## a) Plot the data using pairs command

```{r}
help("birthwt")
names(birthwt)
pairs(birthwt)
GGally::ggpairs(birthwt)
```

## b) Fit a multiple regression model. Summarize the fitted model


```{r}
names(birthwt)
model7 <- lm(bwt~age+lwt+factor(race)+factor(smoke)+ptl+factor(ht)+factor(ui)+ftv, data = birthwt)
tmp <- summary(model7)
kable(tmp$coefficients)

```


## c) Summarize your conclusions. Summarize your fitted model and interpret the coefficients

Checking if linear regression hypothesis are met

```{r}
# 
plot(fitted(model7), residuals(model7), pch=21, col='orange')
abline(h=mean(residuals(model7)), col='seagreen')

# 
plot(birthwt$age, residuals(model7))

tmp

```


It seems that the linear regression assumptions are met: 
- homescedaticity: variance is constant
- mean residuals is 0

The significant factor to predict baby's weight at birth are lwt, race, smoke, ht and ui.



