---
title: "HW5 - "
output: html_notebook
---

```{r include=FALSE}
library(ggplot)
library(MASS)
library(knitr)
```

# Question 1 - Stack loss data

```{r include=FALSE}
# checking the dataset
names(stackloss)
help("stackloss")
```

a) Plot the data

```{r}
pairs(stackloss, font.labels = 3, font.axis = 5, pch = 21)
```


b) Fit a multiple regression model to predict stackloss from the three other variables

Multiple Liner Regression Model: $\beta_0 + \beta_1 * Airflow + \beta_2 * Watertemp + \beta_3 * Acid$

```{r}
model <- lm(stack.loss ~ Air.Flow + Water.Temp + Acid.Conc. , data = stackloss)
summary(model)
```

c) Construct 90% confidence intervals for the coefficients

confidence interval formula: beta +- t_{n-2}(0.05) * std(beta)


```{r}

# using R function
kable(confint(model, level = 0.9), digits = 2, caption = "90% Confidence Intervals for Multiple Linear Regression Coefficients")

# doing it by hand (example with beta0)

n <- nrow(stackloss)
left <- summary(model)$coefficients[1,1] - qt(0.05, df = n-2, lower.tail = FALSE) * summary(model)$coefficients[1,2]
right <- summary(model)$coefficients[1,1] + qt(0.05, df = n-2, lower.tail = FALSE) * summary(model)$coefficients[1,2]
```


d) Construct a 99% prediction interval for a new observation when Airflow=58, Acid=86 and Water Temp=20 (CHECK!!)


```{r}

```




e) Test the null hypothesis H0: beta3 = 0. What is the p-value? What is the conclusion 
if $\alpha=0.10$

the p-value is given by: P(|t| > t_{n-2}(0.05))

```{r}

# compute test statistic t
beta3 <- summary(model)$coefficients[4, 1]
std_beta3 <- summary(model)$coefficients[4, 2]
t <- abs((beta3 - 0) / std_beta3)

# compute p-value
2 * pt(t, df=n-2, lower.tail = FALSE )

summary(model)

```
The p-value for beta3 is 0.3304 > 0.1, so we don't reject H0. This means that 
the acid concentration is not significative enough to be factored in our model.


```{r}
anova(model)
```

