---
title: "HW5 - "
output: html_notebook
---

```{r include=FALSE}
library(ggplot)
library(MASS)
library(knitr)
library(xtable)
```

# Question 1 - Stack loss data

```{r include=FALSE}
# checking the dataset
names(stackloss)
help("stackloss")
```

## a) Plot the data

```{r}
pairs(stackloss, font.labels = 3, font.axis = 5, pch = 21)
```


## b) Fit a multiple regression model to predict stackloss from the three other variables

Multiple Liner Regression Model: $\beta_0 + \beta_1 * Airflow + \beta_2 * Watertemp + \beta_3 * Acid$

```{r}

model <- lm(stack.loss ~ Air.Flow + Water.Temp + Acid.Conc. , data = stackloss)
summary(model)
```

## c) Construct 90% confidence intervals for the coefficients

confidence interval formula: $beta \pm t_{n-2}(0.05) * std(beta)$


```{r}

# using R function
kable(confint(model, level = 0.9), digits = 2, caption = "90% Confidence Intervals for Multiple Linear Regression Coefficients")

# doing it by hand (example with beta0)

n <- nrow(stackloss)
left <- summary(model)$coefficients[1,1] - qt(0.05, df = n-2, lower.tail = FALSE) * summary(model)$coefficients[1,2]
right <- summary(model)$coefficients[1,1] + qt(0.05, df = n-2, lower.tail = FALSE) * summary(model)$coefficients[1,2]

```


## d) Construct a 99% prediction interval for a new observation when Airflow=58, Acid=86 and Water Temp=20 (CHECK!!)


```{r}
names(stackloss)
predict(model, newdata = data.frame(Air.Flow=58, Water.Temp=20, Acid.Conc. = 86), interval = c("none", "confidence", "prediction"), level = 0.95)

predict(model, newdata = data.frame(Air.Flow=58, Water.Temp=20, Acid.Conc. = 86), interval = "prediction", level = 0.95)

x <- predict(model, newdata = data.frame(Air.Flow=58, Water.Temp=20, Acid.Conc. = 86))

# qt(0.025, df = n-2, lower.tail = FALSE)

```


## e) Test the null hypothesis H0: beta3 = 0. What is the p-value? What is the conclusion 
if $\alpha=0.10$

the p-value is given by: $P(|t| > t_{n-2}(0.05))$

```{r}

# compute test statistic t
beta3 <- summary(model)$coefficients[4, 1]
std_beta3 <- summary(model)$coefficients[4, 2]
t <- abs((beta3 - 0) / std_beta3)

# compute p-value
2 * pt(t, df=n-2, lower.tail = FALSE )

summary(model)

```
The p-value for beta3 is 0.3304 > 0.1, so we don't reject H0. This means that 
the acid concentration is not significative enough to be factored in our model.


```{r}
anova(model)
```
# Question 3 - 

## a) Fit the multiple regression in R and summarize the results


```{r}
X <- matrix(c(1,1,1,1,4,3,10,7,5,4,9,10), nrow = 4)
Y <- matrix(c(25, 20, 57, 51), ncol=1)
model2 <- lm(Y ~ X -1) 
model_summary <- summary(model2)$coefficients
row.names(model_summary) <- c("Intercept", "X1", "X2")
kable(model_summary)
```


## b) Construct X^TX and (X^TX)^{-1}

```{r}

t(X) %*% X
solve(t(X) %*% X)

# pretty print with xtable
# print(xtable(t(X) %*% X), tabular.environment = "pmatrix", include.rownames = FALSE, include.colnames = FALSE)

```


## c) Construct $\hat{\beta}$ and confirm the result with R

We know that $\hat{\beta} = (X^TX)^{-1}X^TY$

```{r}
# with formula
solve(t(X) %*% X) %*% t(X) %*% Y

```

We can verify our work using `summary(model2)` and we see that the coefficients 
values are the same

## d) Construct the hat matrix H

We know that $\hat{H} = (X^TX)^{-1}X^T$

```{r}
solve(t(X) %*% X) %*% t(X)
```

## e)  Compute Var($\hat{\beta}$)

$Var(\hat{\beta}) = \sigma^2 (X^TX)^{-1}$

```{r}

```



