---
title: "Chapitre 6 - Time Series Decomposition"
author: "Emulie Chhor"
date: "Feb 27, 2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("TTR")
library("forecast")
library("ggplot2")
library("GGally")
library("lmtest")
library("seasonal")
```

# Load Data

```{r}
births <- scan("http://robjhyndman.com/tsdldata/data/nybirths.dat")
births_ts <- ts(births, start = c(1946, 1))
births_seasons <- ts(births, frequency=12, start = c(1946, 1))
```

# 6.1. Time Series Decomposition

We can decompose a time series into: 
- addictive decomposition
- multiplicative decomposition

# 6.2. Moving Averages: 

- Classical Moving Averages:
- Weighted Moving Averages:

# 6.3. Classical Decomposition

- Additive Decomposition: $\hat{R_t} = y_t - T_t - S_t$
- Multiplicative Decomposition: $\hat{R_t} = y_t / (T_t * S_t)$

```{r}
decompose(births_seasons, type = 'additive') %>% 
  autoplot()

decompose(births_seasons, type = 'multiplicative') %>% 
  autoplot()
```

# 6.4 to 6.6. More Decomposition: X11, SEATS, STL Decompositions

**X11 Decomposition**

```{r}
# 
x11_model <- births_seasons %>%  seas(x11="")
autoplot(births_seasons, series='Data') +
  autolayer(trendcycle(x11_model), series='Trend') +
  autolayer(seasadj(x11_model), series = 'Seasonally Adjusted') + 
  autolayer(births_seasons - x11_model$data[, 'seasonal'], series = 'No Seasonality')

# 
x11_model %>% 
  seasonal() %>% 
  ggseasonplot()

# sub series plots
x11_model %>% 
  seasonal() %>% 
  ggsubseriesplot()
  
```

**SEATS Decomposition**

**STL Decomposition**

Advantages: 
- Handle any type os seasonality, not only monthly and quarterly
- Seasonality can change over time

Disadvantages:
- doesn't handle calendar events automatically

Parameters:
- $ 0 \leq \lambda \leq 1$: from multiplicative decomposition to additive decomposition

```{r}
births_seasons %>% 
  stl(t.window = 13, s.window = 'periodic', robust = TRUE) %>% 
  autoplot()
```

# 6.7. Measuring Strength of a Trend

Measuring the strength of: 
- Trend: $F_t = max(0, 1- \frac{Var(R_t)}{Var(T_t + R_t)})$
- Seasonality: $F_s = max(0, 1- \frac{Var(R_t)}{Var(S_t + R_t)})$

```{r}
R_t <- x11_model$data[, 'final']
S_t <- x11_model$data[, 'seasonal']
T_t <- x11_model$data[, 'trend']

# calculating force of trend and seasonality
F_t <- max(0, 1- var(R_t) / var(R_t + T_t))
F_s <- max(0, 1- var(R_t) / var(R_t + S_t))
```

# 6.8. Forecasting with Decomposition


