---
title: "Chapitre 6 - Time Series Decomposition (Exercices)"
author: "Emulie Chhor"
date: "Feb 27, 2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# install rethinking
library("devtools")
# devtools::install_github("rmcelreath/rethinking", force=TRUE)
# library("rethinking")

# importing packages
library("TTR")
library("forecast")
library("ggplot2")
library("GGally")
library("lmtest")
library("seasonal")
library("tsoutliers")
library("fma")
library("expsmooth")

```

# Q1 - Show that a 3x5 MA is equivalent to a 7-term weighted moving average with weights ...

**What is a 3x5 MA**


**What is a 7-term weighted MA**

**Proving Equivalence**

To show that these two methods are equivalent, we need to show that the weights we
find using 3x5 give us the same answer than when using the weight given by the
7-th term weighted MA given by the system (see chat-gpt for explanation)


```{r}

```

# Q2 - `plastics` dataset (from `fma` package)

**a) Plot time series of sales**

```{r}
# plotting the time series
autoplot(plastics)
```

Trend: it seems that the sales tend to go upward
Seasonality: prices seems to increase the same time of the year (around Christmas)

**b) Use classical multiplicative decomposition to calculate the trend-cycle and seasonal indices**

```{r}
# plotting decomposed time series
plastics %>% 
  decompose(type="additive") %>% 
  autoplot()

# plastics %>%
#   decompose(type="multiplicative") %>%
#   autoplot()
```

**c) Do the results support the graphical interpretation from a)**

Yes

**d) Compute and plot seasonally adjusted data**

We remove seasonality component

```{r}
plastics_adjusted <- plastics - decompose(plastics)$seasonal

# plot adjusted
autoplot(plastics, col='grey', series = 'unadjusted') +
  autolayer(plastics_adjusted, series = 'adjusted')

# plot decomp
plastics_adjusted %>% 
  decompose(type = 'multiplicative') %>% 
  autoplot()
```

**e) change one observation to be an outlier and recompute seasonally adjusted data. What effect has the outlier?**

```{r}
plastics_outliers <- plastics
plastics_outliers[5] <- plastics_outliers[5] + 500
plastics_outliers_adjusted <- plastics_outliers - decompose(plastics_outliers)$seasonal

autoplot(plastics_outliers_adjusted)
platics_outliers %>% 
  decompose(type='multiplicative') %>% 
  autoplot()
```
**f) Does it make any difference if the outlier is near the end rather than in the middle**

```{r}
length(plastics)

# middle
plastics_outliers_middle <- plastics
plastics_outliers_middle[30] <- plastics_outliers_middle[5] + 500
plastics_outliers_middle_adjusted <- plastics_outliers_middle - decompose(plastics_outliers_middle)$seasonal

autoplot(plastics_outliers_middle_adjusted)
platics_outliers %>% 
  decompose(type='multiplicative') %>% 
  autoplot()

# end
plastics_outliers_end <- plastics
plastics_outliers_end[58] <- plastics_outliers_end[58] + 500
plastics_outliers_end_adjusted <- plastics_outliers_end - decompose(plastics_outliers_end)$seasonal

autoplot(plastics_outliers_end)
platics_outliers %>% 
  decompose(type='multiplicative') %>% 
  autoplot()
```

Yes it does! It seems that when the outlier is at the end of the time series, 
adjusting it doesn't have much effect vs when the outlier is in the middle 
of the time series.

# Q3 - X11 Decomposition - `dj` (Dow Jones)

```{r}
# 2.10. computing daily changes, plot and ACF
ddj <- diff(dj)

autoplot(dj) 
autoplot(ddj)
acf(ddj)

# x11 decomposition on 
dj_cleaned <- ts(dj, start="1946", frequency = 12)
dj_x11_decomp <- dj_cleaned %>% seas(x11="")

autoplot(dj_x11_decomp)

# finding outliers using tso()
# dj_decomp <- decompose(dj_cleaned, type = "multiplicative")
# outliers <- tso(dj_decomp, types = c("AO", "LS", "TC"), maxit = 10)

```

# Q4 - 

**a) Results of the decomposition**

By looking at the decomposition, we see that from February 1978 to August 1995, 
the number of civilian working keeps increasing over the years. There also
seems to a seasonal component to the number of workers in the workforce: the
number of workers seem to be the highest in March and December and the lowest 
in January and August. 

**b) Is the recession of 1991/1992 visible in the estimated components**

Yes, the recession is visible in the estimated components: we see that 
the remainder around 1991/1992 is a lot less than the expected values.

# Q5 - `cangas` dataset from `expsmooth`

**a) Plot the data. Effect of seasonality over time**

```{r}
autoplot(cangas)
ggsubseriesplot(cangas)
ggseasonplot(cangas)
```
It seems that in the months where it is colder (ie during the winter), Canadian gas consumption is a higher than in months when temperature is warmer.

**b) STL Decomposition**

```{r}
#
cangas %>% decompose(type='multiplicative')

# stl decomposition
cangas_stl_decomp <- stl(cangas, s.window = 12)
autoplot(cangas_stl_decomp)
```

**c) Compare with SEATS and X11 decomposition**

```{r}
# SEATS decomposition
cangas_seats_decomp <- cangas %>% 
  seas()
autoplot(cangas_seats_decomp)

# x11 decomp
cangas_x11_decomp <- cangas %>% 
  seas(x11="")
autoplot(cangas_x11_decomp)

```

The STL decomposition is a bit smoother than the x11 and SEATS decompositions.

# Q6 - `bricksq` dataset

**a) STL decomposition to calculate trend-cycle and seasonal indices**

```{r}
bricksq_stl <- stl(bricksq, s.window = 4)
autoplot(bricksq_stl)
```

**b) Plot seasonally adjusted data**

```{r}
bricks_adjusted_stl <- bricksq - bricksq_stl$time.series[,1]
bricks_adjusted_decomp <- bricksq - decompose(bricksq)$seasonal
autoplot(bricksq, series = "Normal", col='grey') + 
  autolayer(bricks_adjusted_stl, series = "Adjusted with decomp") +
  autolayer(bricks_adjusted_decomp, series = "Adjusted with STL")
```

**c) Use naive method to forecast seasonally adjusted data**

```{r}
fit <- stl(bricksq, t.window=12, s.window="periodic",
  robust=TRUE)
fit %>% forecast(method="naive") %>%
  autoplot()
```

**d) Use stlf() to reseasonalise the results**

```{r}

```


**e) Do the residuals look uncorrelated?**

**f) Repeat with a robust STL decomposition. Does it make a difference?**

**g) Compare forecasts from stlf() with those from snaive() asing a test set. Which is better**

# Q7 - `writing` dataset


# Q8 - `fancy` dataset










